---
title: "pkgsimil"
output:
  md_document:
    variant: gfm

  rmarkdown::html_vignette:
    self_contained: no
---

<!-- badges: start -->

[![R build
status](https://github.com/ropensci-review-tools/pkgsimil/workflows/R-CMD-check/badge.svg)](https://github.com/ropensci-review-tools/pkgsimil/actions?query=workflow%3AR-CMD-check)
[![codecov](https://codecov.io/gh/ropensci-review-tools/pkgsimil/branch/main/graph/badge.svg)](https://app.codecov.io/gh/ropensci-review-tools/pkgsimil)
[![Project Status:
WIP](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
<!-- badges: end -->

# pkgsimil

Similarity metrics between R packages, functions, or any code, based on
similarity of code, documentation, or both.

If the package has not yet been installed, the following line needs to be run:

```{r eval = FALSE}
remotes::install_github ("ropensci-review-tools/pkgsimil")
```

The package can then be loaded for us with:

```{r library}
library (pkgsimil)
```

The package measures similarity from Large Language Model (LLM) embeddings and
requires a locally-running instance of [ollama](https://ollama.com), as
described in the following sub-section.


## Setting up the LLM embeddings

The LLM embeddings are extracted from a locally-running instance of
[ollama](https://ollama.com). That means you need to download and install
ollama on your own computer in order to use this package. Once downloaded,
ollama can be started by calling `ollama serve`. The particular models used to
extract the embeddings will be automatically downloaded by this package if
needed, or you can do this manually by running the following two commands (in a
system console; not in R):

``` bash
ollama pull jina/jina-embeddings-v2-base-en
ollama pull ordis/jina-embeddings-v2-base-code
```

You'll likely need to wait up to half an hour or more for the models to
download before proceeding.

## Using the `pkgsimil` package

The package has two main functions:

- `pkgsimil_similar_pkgs()` to find similar rOpenSci packages based input as
either a local path to an entire package, or as a single descriptive text
string; and
- `pkgsimil_similar_fns()` to find similar functions from rOpenSci packages based
on descriptive text input.

The following code demonstrates how these functions work, first with two
demonstrates of finding packages:

```{r pkgs1}
input <- "
Packages for analysing evolutionary trees, with a particular focus
on visualising inter-relationships among distinct trees.
"
pkgsimil_similar_pkgs (input)
```

```{r pkgs2}
input <- "Download global-scale spatial data"
pkgsimil_similar_pkgs (input)
```


And then two demonstrates of finding functions from rOpenSci packages:

```{r fns1}
input <- "A function to label a set of geographic coordinates"
pkgsimil_similar_fns (input)
```
```{r fns2}
input <- "Identify genetic sequences matching a given input fragment"
pkgsimil_similar_fns (input)
```

## Prior Art

- The [`utils::RSiteSearch()`
function](https://stat.ethz.ch/R-manual/R-devel/library/utils/html/RSiteSearch.html).
- The [`sos` package](https://github.com/sbgraves237/sos) that queries the
"RSiteSearch" database.
