---
title: "The pkgmatch package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{pkgmatch}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set (
    collapse = TRUE,
    comment = "#>"
)
```

The "pkgmatch" package is a search and matching engine for R packages. It finds
the best-matching R packages to an input of either a text description, or a
local path to an R package. `pkgmatch` was developed to enable rOpenSci to
identify similar packages to each new package submitted for [our software
peer-review scheme](https://ropensci.org/software-review/). By default, matches
are found from [rOpenSci's own package suite](https://ropensci.org/packages/),
but it is also possible to find matches from all [packages currently on
CRAN](https://cran.r-project.org).

## What does the package do?

What the package does is best understood by example, starting with loading the package.

```{r library}
library (pkgmatch)
```

Then match packages to an input string:

```{r match-text-1-fakey, eval = FALSE}
input <- "Download and visualize open spatial data"
pkgmatch_similar_pkgs (input)
```

```{r redef-sim-pkgs1, eval = TRUE, echo = FALSE}
c ("GSODR", "landscapetools", "rentrez", "onekp", "geojsonio")
```

The following lines find equivalent matches against all packages currently on
CRAN:

```{r match-text-2-cran-fakey, eval = FALSE}
pkgmatch_similar_pkgs (input, corpus = "cran")
```

```{r redef-sim-pkgs2, eval = TRUE, echo = FALSE}
c ("RImageJROI", "OCSdata", "ggspatial", "ppcSpatial", "rleafmap")
```

As said, the package also accepts as input a path to a local R package. This
code downloads a "tarball" (`.tar.gz` file) from CRAN and finds matching
rOpenSci packages:

```{r odbc-ropensci-match-fakey, eval = FALSE}
u <- "https://cran.r-project.org/src/contrib/odbc_1.5.0.tar.gz"
destfile <- file.path (tempdir (), basename (u))
download.file (u, destfile = destfile)
pkgmatch_similar_pkgs (destfile)
```

```{r odbc-ropensci-match, eval = TRUE, echo = FALSE}
list (
    text = c ("taxadb", "pdftools", "drake", "pathviewr", "baRulho"),
    code = c ("gistr", "pkgstats", "melt", "aorsf", "postdoc")
)
```

We of course expect the best matches against CRAN packages to include that
package itself:

```{r odbc-cran-match-fakey, eval = FALSE}
pkgmatch_similar_pkgs (destfile, corpus = "cran")
```

```{r odbc-cran-match, echo = FALSE, eval = TRUE}
list (
    text = c ("lutz", "wesanderson", "odbc", "implyr", "MM"),
    code = c ("PointFore", "odbc", "pkgload", "phrases", "chemdeg")
)
```

which they indeed do, although the package is not ranked first either in terms
of text matches or code matches. As explained in the documentation, the
`pkgmatch_similar_pkgs()` function ranks final results by combining several
distinct results, primarily from Large Language Model (LLM) embeddings, as well
as from [more conventional document token-frequency
analyses](https://en.wikipedia.org/wiki/Okapi_BM25). The function includes an
additional `llm_proportion` parameter which can be used to weight the relative
contributions of these different inputs. Results from the LLM component are:

```{r odbc-cran-match-llm-fakey, eval = FALSE}
pkgmatch_similar_pkgs (destfile, corpus = "cran", llm_proportion = 1)
```
```{r odbc-cran-match-llm, echo = FALSE, eval = TRUE}
list (
    text = c ("wesanderson", "lutz", "MM", "growthmodels", "lunar"),
    code = c ("PointFore", "phrases", "chemdeg", "ShiftConvolvePoibin", "date")
)
```

And the package does not even appear in terms of similarity in either text or
code. Results from other other component, comparing relative token frequencies
with all CRAN packages, including frequencies of code tokens, are:

```{r odbc-cran-match-bm25-fakey, eval = FALSE}
pkgmatch_similar_pkgs (destfile, corpus = "cran", llm_proportion = 1)
```
```{r odbc-cran-match-bm25, echo = FALSE, eval = TRUE}
list (
    text = c ("odbc", "implyr", "DatabaseConnector", "sparklyr", "gbifdb"),
    code = c ("odbc", "pkgload", "httr2", "gganimate", "gargle")
)
```

And those analyses correctly identify the package as the most similar in both
ways.
