---
title: "Why local LLMs?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Why local LLMs?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set (
    collapse = TRUE,
    comment = "#>"
)
```

The "pksimil" package uses Large Language Models (LLMs) to assess relationships
between R packages. Software which relies on LLMs commonly accesses them
through Application Programming Interfaces (APIs) provided by external
organisations such as [mistral.ai](https://mistral.ai),
[jina.ai](https://jina.ai), or a host of alternative providers.
Inputs, generally in text form, are sent to the external service which then
responds in some specified form, such as text completions.

Accessing LLMs through APIs has several notable drawbacks, including:

- There is no guarantee that the API will continue to be available in the
future, or that processes used to generate responses will remain stable and
reproducible.
- Most APIs cost money. These costs must generally be borne by the users of
software.
- Data submitted to such APIs is generally used by the organizations providing
them to train and refine models, so privacy-protecting use is generally not
possible.

In spite of those drawbacks, building software around external APIs offers the
two key advantages of:

- Being easier to develop, as the external API will generally take care of much
of the processing that might otherwise have to be written and executed locally;
and
- Being able to access the latest and biggest and fastest models which are
generally only available in the form of external APIs.

The "pkgsimil" package interfaces with LLMs exclusively through a local server
provided by [the "ollama" software](https://ollama.com). The "pkgsimil" package
relies on two openly published LLM models, both from
[jina.ai](https://jina.ai):

- [jina-embeddings-v2-base-en](https://huggingface.co/jinaai/jina-embeddings-v2-base-en)
- [jina-embeddings-v2-base-code](https://huggingface.co/jinaai/jina-embeddings-v2-base-code)

```{r setup}
library (pkgsimil)
```
